Sequential algorithms to split and merge terabyte-scale 3D images

### 2017-06-12 paper.tex
gitignore and Figure 1

As detailed in Algorithm~\ref{algo:cluster-reads}, cluster reads load
multiple blocks in memory, concatenate them in a buffer and write the
buffer as sequentially as possible. The buffer is a slightly complex
data structure, for instance an associative array or a Python
dictionary, capable of storing multiple disjoint sequences of
contiguous bytes without having to allocate memory for the bytes
between such sequences. Implementation details about the buffer are
given later in the document.

\todo{a figure showing the buffer and how it
  fillus up would help}.
%\note{More on defragmentation might be
%  needed here.}
\begin{algorithm}[h]
  \caption{Buffered merging from blocks using cluster reads}
  \label{algo:cluster-reads}
  \begin{algorithmic}
    \STATE list = sort blocks by increasing (x,y,z) values
    \STATE initialize buffer
    \STATE index = 0
    \WHILE{index\textless n}
      \WHILE{sizeof(buffer)\textless m}
        \STATE block = list[index]
        \STATE read block and insert it in buffer
        \STATE index += 1
        \ENDWHILE
       \STATE write buffer in reconstructed image
       \STATE clear buffer
    \ENDWHILE
  \end{algorithmic}
\end{algorithm}

The number of seeks performed by cluster reads depends on the way that
blocks loaded in memory arrange in the reconstructed image. In the
best case, blocks in memory completely cover slices in the
reconstructed image. In this case, slices of the reconstructed images
can be written without seeking at all. If the blocks in memory
completely cover rows in the reconstructed image but do not cover
entire slices, then the blocks are written with $d$ seeks, one per
slice. And if the blocks do not cover entire rows, then they are
written with $d^2$ seeks, one per row and per slice.  In conclusion:
\begin{equation}
  N_\mathrm{cluster\_reads}= n + \ceil*{\frac{bR}{m}}f(m,n,R)
\end{equation}
where:
\begin{equation}
  f(m,n,R)=
  \begin{cases}
    d^2 & $if there is an incomplete row$\\
    d   & $if there is an incomplete slice but not incomplete row$\\
    1   & $otherwise$
\end{cases}
\end{equation}

>>>>

As detailed in Algorithm~\ref{algo:cluster-reads}, cluster reads load
multiple blocks in memory, concatenate them in a buffer and write the
buffer sequentially in the reconstructed image. The buffer is a slightly complex
data structure, for instance an associative array or a Python
dictionary, capable of storing multiple disjoint sequences of
contiguous bytes without having to allocate memory for the bytes
between such sequences. Implementation details about the buffer are
given later in the document.

\begin{figure}
\centering
\def\svgwidth{0.3\columnwidth}
\input{figures/no-seek.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/d-seeks.pdf_tex}
\def\svgwidth{0.3\columnwidth}
\input{figures/d2-seeks.pdf_tex}
\caption{Left: complete slices and rows are reconstructed in memory:
  no seeking is required. Center: complete rows but incomplete slices
  are reconstructed: $d$ seeks are required. Right: incomplete rows
  are reconstructed: $d^2$ seeks are required.}
\label{fig:cluster-reads}
\end{figure}

\todo{a figure showing the buffer and how it
  fill up would help}.
%\note{More on defragmentation might be
%  needed here.}
\begin{algorithm}[h]
  \caption{Buffered merging of blocks with cluster reads}
  \label{algo:cluster-reads}
  \begin{algorithmic}
    \STATE list = sort blocks by increasing (x,y,z) values
    \STATE initialize buffer
    \STATE index = 0
    \WHILE{index\textless n}
      \WHILE{sizeof(buffer)\textless m}
        \STATE block = list[index]
        \STATE read block and insert it in buffer
        \STATE index += 1
        \ENDWHILE
       \STATE write buffer in reconstructed image
       \STATE clear buffer
    \ENDWHILE
  \end{algorithmic}
\end{algorithm}

The number of seeks performed by cluster reads depends on how blocks
loaded in memory arrange in the reconstructed image. In the best case,
complete contiguous slices of the reconstructed image can be assembled
in memory and written without seeking at all. In the worst case, some
rows in the reconstructed image are not covered by blocks loaded in
memory: $d^2$ seeks are then required, one for every partial row in
every partial slice. In the intermediary case, rows are complete but
some slices can only be partially reconstructed: $d$ seeks are then
required. In summary:
\begin{equation}
  N_\mathrm{cluster\_reads}= n + \ceil*{\frac{bR}{m}}f(m,n,R)
\end{equation}
where:
\begin{equation}
  f(m,n,R)=
  \begin{cases}
    d^2 & $if there is an incomplete row$\\
    d   & $if there is an incomplete slice but not incomplete row$\\
    1   & $otherwise$
\end{cases}
\end{equation}
